---
author:
    name: "Angelo Monroy"
date: today
title: "Personal CGM Data EDA"
format: html
---

# Intro

Performing EDA of personal CGM data for use in hidden-instability.

```{r}
#| label: libraries
#| message: false

library(readr)
library(dplyr)
library(tidyr)

# interactive source
# source("R/helpers.R")
# render source
source("../R/helpers.R")
```

# Data

Data was sourced from Dexcom Clarity app January 23, 2026.

```{r}
#| label: load-raw-local
#| message: false

# First CSV in data/raw (e.g. Clarity export)
raw_path <- list.files("data/raw", pattern = "\\.csv$", full.names = TRUE)
if (length(raw_path) == 0) stop("No CSV found in data/raw")
df <- read_csv(raw_path[1], show_col_types = FALSE)
problems(df)
```

CSV had 17 rows missing NA for last column. Manually fixed data and reread csv.

```{r}
#| label: reload-raw-local
#| message: false

df <- read_csv(raw_path[1], show_col_types = FALSE)
```

```{r}
#| label: tidy-dfs

meals <- df |>
    filter(`Event Type` == "Meal") |>
    drop_empty_columns()

cgm <- df |>
    filter(`Event Type` == "EGV") |>
    drop_empty_columns()

old_cgm_colnames <- colnames(cgm)
colnames(cgm) <- c("index", "ts", "type", "subtype", "device", "egv", "device_tick", "device_id")



calibrations <- df |>
    filter(`Event Type` == "Calibration") |>
    drop_empty_columns()
```

# EDA

```{r}
#| label: basic-sum

data_completion <- (nrow(cgm)/1440*5) / as.numeric(max(cgm$ts) - min(cgm$ts))

```

## Summary

* Time Period: `{r} format(min(cgm$ts), "%B %d, %Y")` - `{r} format(max(cgm$ts), "%B %d, %Y")`
* Device: `{r} unique(cgm$device)`
* Data Availability: `{r} data_completion`

## Device Metrics

```{r}
#| label: device metrics

nrow(calibrations)

```

## Behavioral Metrics

* There were 