---
author:
    name: "Angelo Monroy"
date: today
title: "Personal CGM Data EDA"
format: html
---

# Intro

Performing EDA of personal CGM data for use in hidden-instability. Dexcom CGM G7 is an FDA-approved medical device that continuously monitors glucose via measuring of interstitial fluid. It lasts for 10 days + 12 hour grace period.

```{r}
#| label: libraries
#| message: false

library(readr)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(reticulate)

# interactive source
# source("R/helpers.R")
# render source
source("../R/helpers.R")
```

# Data

Data was sourced from Dexcom Clarity app January 23, 2026.

## R

```{r}
#| label: load-raw-local
#| message: false

# First CSV in data/raw (e.g. Clarity export)
raw_path <- list.files("data/raw", pattern = "\\.csv$", full.names = TRUE)
if (length(raw_path) == 0) stop("No CSV found in data/raw")
df <- read_csv(raw_path[1], show_col_types = FALSE)
problems(df)
```

CSV had 17 rows missing NA for last column. Manually fixed data and reread csv.

```{r}
#| label: reload-raw-local
#| message: false

df <- read_csv(raw_path[1], show_col_types = FALSE)
```

```{r}
#| label: tidy-dfs

meals <- df |>
    filter(`Event Type` == "Meal") |>
    drop_empty_columns()

cgm <- df |>
    filter(`Event Type` == "EGV") |>
    drop_empty_columns()

old_cgm_colnames <- colnames(cgm)
colnames(cgm) <- c("index", "ts", "type", "subtype", "device", "egv", "device_tick", "device_id")



calibrations <- df |>
    filter(`Event Type` == "Calibration") |>
    drop_empty_columns()
```

## Python

Reload dataset with Python for glycemic analysis

```{python}
#| label: r-df-to-python
import sys
from pathlib import Path

import numpy as np
import pandas as pd

# Ensure the project root (one level up from notebooks/) is on the Python path
root = Path.cwd()

# Load the same raw CSV used in the R chunks
raw_files = list((root / "data" / "raw").glob("*.csv"))
if not raw_files:
    raise FileNotFoundError("No CSV found in ../data/raw")

df_py = pd.read_csv(raw_files[0])

# Drop columns that are entirely NA (all missing), similar to drop_empty_columns()
df_py = df_py.dropna(axis=1, how="all")

# Split into EGV (CGM), meals, and calibrations, mirroring the R data frames
meals_py = df_py[df_py["Event Type"] == "Meal"].copy()
cgm_py = df_py[df_py["Event Type"] == "EGV"].copy()
calibrations_py = df_py[df_py["Event Type"] == "Calibration"].copy()

# Drop any all-NA columns post-filter as well
meals_py = meals_py.dropna(axis=1, how="all")
cgm_py = cgm_py.dropna(axis=1, how="all")
calibrations_py = calibrations_py.dropna(axis=1, how="all")

# Align CGM column names with the R pipeline when the width matches
if cgm_py.shape[1] == 8:
    cgm_py.columns = [
        "index",
        "ts",
        "type",
        "subtype",
        "device",
        "egv",
        "device_tick",
        "device_id",
    ]

# render source
from src import (
    compute_TIR,
    compute_TBR,
    compute_TAR,
    compute_GMI,
    compute_summary_metrics,
)
```

## By Hour Summarizing

```{python}

#| label: hourly-metrics
import importlib
import src.hourly

# Reload so edits to src/hourly.py take effect without restarting
importlib.reload(src.hourly)
from src.hourly import (
    make_hour_of_day_metrics,
    make_hour_dow_metrics,
)

hour_of_day_metrics_py = make_hour_of_day_metrics(cgm_py)
hour_dow_metrics_py = make_hour_dow_metrics(cgm_py)

hour_of_day_metrics_py.head()
hour_dow_metrics_py.head()
```

# EDA

```{r}
#| label: basic-sum

data_completion <- (nrow(cgm)/1440*5) / as.numeric(max(cgm$ts) - min(cgm$ts))

```

## Summary

* Time Period: `{r} format(min(cgm$ts), "%B %d, %Y")` - `{r} format(max(cgm$ts), "%B %d, %Y")`
* Device: `{r} unique(cgm$device)`
* Data Availability: `{r} pct(data_completion, 1)`
* CGMs Used in Period: `{r length(unique(cgm$device_id))}`
* Manual BGM Calibrations: `{r nrow(calibrations)}`
* Meal Entries: `{r nrow(meals)}`

## Visualizations

```{r}

hourly_df <- as.data.frame(reticulate::py_to_r(py$hour_of_day_metrics_py))

hourly_df$hour <- as.integer(hourly_df$hour)
hourly_df$hour_label <- format(
  as.POSIXct(sprintf("2000-01-01 %02d:00:00", hourly_df$hour),
             format = "%Y-%m-%d %H:%M:%S",
             tz = "America/Los_Angeles"),
  "%I:%M %p"
)

rects <- data.frame(
  ystart = c(40, 70, 180),
  yend   = c(70, 180, 300),
  col    = c("Low", "In Range", "High")
)

ggplot(hourly_df, aes(x = factor(hour, labels = hour_label), y = mean)) +
  geom_rect(
    data = rects,
    aes(ymin = ystart, ymax = yend, xmin = -Inf, xmax = Inf, fill = col),
    inherit.aes = FALSE,
    alpha = 0.4
  ) +
  geom_line(group = 1) +
  scale_y_continuous(
    limits = c(40, 300),
    breaks = c(40, 70, 100, 120, 150, 200, 250, 300)
  ) +
  labs(
    x = "Hour of Day",
    y = "Mean Blood Glucose Mg/dL",
    title = "Mean Glucose by Hour"
  ) +
   theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

## Glycemic Analysis

```{python}
#| label: glyc-metrics
from src import (
    compute_TIR,
    compute_TBR,
    compute_TAR,
    compute_GMI,
    compute_summary_metrics,
)

# Use the CGM glucose values as a 1D NumPy array
glucose = cgm_py["egv"].to_numpy(dtype=float)

metrics_unmasked = {
    "TIR": compute_TIR(glucose),
    "TBR": compute_TBR(glucose),
    "TAR": compute_TAR(glucose),
    "GMI": compute_GMI(glucose),
    "summary": compute_summary_metrics(glucose),
}

metrics_unmasked
```

# References

Cichosz SL, Xylander AAP. A Conditional Generative Adversarial Network for Synthesis of Continuous Glucose Monitoring Signals. J Diabetes Sci Technol. 2021 May 30:19322968211014255. doi: 10.1177/19322968211014255. Epub ahead of print. PMID: 34056935